---
title: "experiments2022"
author: "mira"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(dplyr)
library(ggplot2)
library(viridis)
library(sp)
library(fields)
library(gstat)
library(gear)

```

# spatial conformal prediction: univariate 

## scp prediction 
demonstration with univariate data. krigging based point prediction and conformal based confidence interval.
ref
Barber, R. F., Candes, E. J., Ramdas, A., & Tibshirani, R. J. (2022). Conformal prediction beyond exchangeability. arXiv preprint arXiv:2202.13415.

Mao, H., Martin, R., & Reich, B. (2020). Valid model-free spatial prediction. arXiv preprint arXiv:2006.15640.https://github.com/mhuiying/scp

Mattern cov func, Bessel func $K_k(.)$
$\rho(u;\phi,\kappa)=\{2^{\kappa-1}\Gamma(\kappa)\}^{-1}(\frac{u}{\phi})^{\kappa} K_k(\frac{u}{\phi})$

```{r utils covariance matrix,include=TRUE, eval=FALSE}
# Matern covariance fuction 
# matern, https://github.com/cran/geoR/blob/master/R/corcov.R
# u distance, vec matrix or array of the distances between point pairs
# phi the range parameter
# kappa the smoothness parameter 
# $\rho(u;\phi,\kappa)=\{2^{\kappa-1}\Gamma(\kappa)\}^{-1}(\frac{u}{\phi})^{\kappa} K_k(\frac{u}{\phi})$
matern <- function (u, phi, kappa) {
  if(is.vector(u)) names(u) <- NULL
  if(is.matrix(u)) dimnames(u) <- list(NULL, NULL)
  uphi <- u/phi
  uphi <- ifelse(u > 0,
                 (((2^(-(kappa-1)))/ifelse(0, Inf,gamma(kappa))) *
                  (uphi^kappa) *
                  besselK(x=uphi, nu=kappa)), 1)    
  uphi[u > 600*phi] <- 0 
  return(uphi)
}
# d a numeric distance, a vector or matrix
# theta parameter [Nugget, Sill, Range, Smoothness]
mat_cov <- function(d,theta){
  theta[1]*diag(nrow(d)) + theta[2]*matern(d,theta[3],theta[4])
}

```


```{r test setup,include=TRUE, eval=FALSE}
# test setup
set.seed(123)
N <- 11
n <- N^2
S <- seq(0,1,length=N) # 11
s <- expand.grid(S,S) # 121
d <- as.matrix(dist(s)) # 121*121

theta <- c(0,3,0.1,0.7)
names(theta) <- c("Nugget","Sill","Range","Smoothness")
C <- mat_cov(d,theta) # 121*121

X <- t(chol(C))%*%rnorm(n) # 121*1
Y <- X^3 + rnorm(n) # 121*1
plot(Y~X)


```


```{r Estimate theta,include=TRUE, eval=FALSE}
# estimate parameters for covariance via variogram fitting
# variogram est with fields::vgram which returns a vgram object

# s spatial locations
# Y 
# dists breakpoints for variogram bins
# return Matern covariance parameters theta

get_theta <- function(s, Y, dists = NULL) {
  # variogram estimation (not binned?), 
  s<- as.matrix(s)
  max_range=max(dist(s))
  vg <- fields::vgram(s, Y, dmax=max_range, breaks = dists,type="variogram") 
  # vgran() ouput d, vgram ...
  
  wsse <- function(logtheta, d, v, max_range) {
    # theta to be optimised, 
    # d a vector of distances
    # v a vector of estimated variogram values at distance d
    # n number of pairs in each bin for binned variogram
    theta <- exp(logtheta)
    theta[4] <- max_range*pnorm(logtheta[4])
    m  <- theta[1] + theta[2] - theta[2]*matern(d,theta[3],theta[4])
    out <- sum((v - m)^2)
    return(out)}
  
  init <- c(0.5 * var(Y, na.rm = TRUE), 0.5 * var(Y, na.rm = TRUE),
            quantile(vg$d,0.5), 0.5)
  
  fit <- optim(log(init), wsse, d = vg$d, v = vg$vgram, max_range=max_range) 
  
  theta <- exp(fit$par)
  theta[4] <- max_range*pnorm(fit$par[4])
  names(theta) <- c("Nugget", "Sill", "Range", "Smoothness")
  return(theta)
  }
```


```{r,include=TRUE, eval=FALSE}
######################## test get_theta
# Estimate spatial covariance parameters
# oracle: theta = c(0,  3,  0.1,  0.7)

bins <- seq(0.01,0.2,0.01) # 20 bins
thetaHat <- get_theta(s,Y,dists=bins)

thetaHat # 65.8,  29.2,  2.46, 0.000086

```


```{r select eta kernel bandwidth,include=TRUE, eval=FALSE}
# calculate interval score
# interval lower bound l, upper bound u, data x, significance level alpha
# return a value of interval score
interval_score <- function(l, u, x, alpha = 0.05){
  width <- u - l
  sl <- 2 / alpha * (l - x) * (x < l)
  su <- 2 / alpha * (x - u) * (x > u)
  width + sl + su
}

# select eta the tuning parameter for scp, kernel bandwidth for weight schema
# s
# Y
# eta_cand a vector of canditate eta values, default to null
# return optimum eta based on minimizing empirical interval score

select_eta <- function(s,Y,eta_cand = NULL){
  bins <- seq(0.01,0.2,0.01)
  thetaHat <- get_theta(s,Y,dists=bins)
  min_d <- min(dist(s))
  eta_cand <- seq(3*min_d, 20*min_d, min_d)
  # random sample of prediction point
  valid_id <- sample(1:nrow(s),1)
  # compute and store intScore for candidate eta values
  intScore <- c()
  for(i in 1:length(eta_cand)){
    PIs <- scp(s0=s[valid_id,],s=s[-valid_id,],Y=Y[-valid_id],thetaHat=thetaHat,eta=eta_cand[i],m=20,precision=0.1)
    intScore <- c(intScore, interval_score(PIs[1], PIs[2], Y[valid_id]))
  }
  opt_eta <- eta_cand[which.min(intScore)]
  # plot 
  plot(eta_cand, intScore, type = "b", xlab = "eta", ylab = "interval score")
  abline(v = opt_eta, col = "red", lty=2)
  return(opt_eta)
}
```


```{r, eval=FALSE}
############################## test select_eta

test.eta <- select_eta(s,Y,eta_cand = NULL) # not varying by eta_cand ???
test.eta # 0.3

```


```{r prime,include=TRUE, eval=FALSE}
# Transform the input for local SCP prediction
# s0 prediction location 1x2 
# s n x 2 matrix or data.frame
# Y n dim
# eta kernel bandwidth. Defauls to Inf equal weight on m points.
# m number of nearest locations to use for prediction.

.prime <- function(s0,s,Y,eta,m){
  # Determine these surrounding m points for prediction， LSCP
  dist <- sqrt( (as.numeric(s0[1])-s[,1])^2 + (as.numeric(s0[2])-s[,2])^2 )
  these <- order(dist)[1:m] # vector of location index
  # subset selected points
  s_aug <- rbind(s0, s[these,])
  Y_aug <- c(NA, Y[these])
  d_aug <- as.matrix(dist(s_aug))
  # calculate weights w
  M <- length(these)
  if(eta==Inf){
    w <- rep(1/(M+1),M+1) # equal weights to all m points
  }else{
    w <- d_aug[1,-1]/eta
    w <- c(1, exp(-0.5*w^2)) # exponential decay 
    w <- w/sum(w)
  }
  # output
  s_aug <<- s_aug
  Y_aug <<- Y_aug
  w_aug <<- w
  d_aug <<- d_aug
}
```


```{r compute_UVW,include=TRUE, eval=FALSE}
# compute U, V, W when delta_i = (Y_i - Yhat_i)^2
# Q inverse of the covariance matrix
# Y 
# return a list of U, V, W values
.compute_UVW2 <- function(Q,Y){
  Ysub <- Y[-1]
  Qrow <- Q[1,-1]
  Qsub <- Q[-1,-1]
  dQsub <- diag(Qsub)
  Yhat <- sum(Qrow*Ysub)/Q[1,1]
  Ytilde <- Qsub%*%Ysub/dQsub
  U <- Ytilde^2-Yhat^2
  V <- Ytilde*Qrow/dQsub - Yhat
  W <- Qrow*Qrow/(dQsub*dQsub)-1
  out <- list(U=c(0,U),V=c(0,2*V),W=c(0,W))
  return(out)}
```


```{r plausibility,include=TRUE, eval=FALSE}
# Fast plausibility_contour calculation when pred_fun = Krige_fun and dfun = residual^2 and weights are provided.
# s0 prediction location
# s n by d matrix or data-frame with d coordinates of the n locations.
# Y 
# w weights for the non-conformity measures.
# Q n+1 by n+1 inverse of estimated covariance matrix
# alpha significance level. Defaults to 0.05.
# return A vector of lower/ upper bounds of the conformal prediction interval.

.plausibility_contour <- function(Y_aug,w_aug,Q){

  UVW <- .compute_UVW2(Q,Y_aug)
  U <- UVW$U[-1]
  V <- UVW$V[-1]
  W <- UVW$W[-1]

  low <- (-V + sqrt(V^2 - 4*U*W))/(2*W)
  upp <- (-V - sqrt(V^2 - 4*U*W))/(2*W)

  w  <- w_aug[-1]
  lu_w  <- c(w,w)
  lu_lab <- c(rep(1, length(low)), rep(-1, length(upp)))
  lu_w   <- lu_w[order(c(low, upp))]
  lu_lab <- lu_lab[order(c(low, upp))]

  y0 <- sort(c(low, upp))
  p_y <- cumsum(lu_w*lu_lab)

  return(data.frame(Y_cand=y0,p_y=p_y))
}
```


```{r, eval=FALSE}
############################## test plausibility

.prime(s0,s,Y,eta=Inf,m=20) 
Q.test <- solve(mat_cov(d_aug,thetaHat))
p_df  <- .plausibility_contour(Y_aug,w_aug,Q.test)
plot(p_df) # Y_cand narrowly centered around 0.5

.prime(s0,s,Y,eta=0.5,m=20) 
Q.test <- solve(mat_cov(d_aug,thetaHat))
p_df  <- .plausibility_contour(Y_aug,w_aug,Q.test)
plot(p_df) # also works

```


```{r krig prediction func,include=TRUE, eval=FALSE}
# Kriging prediction function for pred_fun, point prediction at s0 given s and Y
# s0 prediction location, a numeric vector with length = 2.
# s n x 2 matrix or data.frame.length(s0)= ncol(s)
# Y n dim length(Y)= nrow(s)
# return a value of point prediction for Y(s0);

krige_pred <- function(s0,s,Y){
  bins <- seq(0.01,0.2,0.01) # bins chosen how???
  thetaHat <- get_theta(s,Y,dists=bins)
  idx <- which(s[,1]==s0[1] & s[,2]==s0[2])
  if(length(idx) > 0){s = s[-idx,]; Y = Y[-idx]}
  # distance matrix
  s_aug <- rbind(s0,s)
  d_aug <- as.matrix(dist(s_aug))
  # inverse covariance matrix
  Q <- solve(mat_cov(d_aug,thetaHat))
  # Kriging prediction
  yhat  <- as.numeric(-Q[1,-1]%*%Y/Q[1,1])
  sd <- 1/sqrt(Q[1,1])
  return(yhat)
}
```


```{r, eval=FALSE}
############################## test kriging
s0 <- c(0.5, 0.5)

idx <- which(s[,1]==s0[1] & s[,2]==s0[2]) # 61
Y[61,] # 1.799352

pred.test <- krige_pred(s0=s0,s=s[-idx,],Y=Y[-idx])
# 0.0138

```


```{r Spatial conformal prediction scp,include=TRUE, eval=FALSE}
# Spatial conformal prediction at single input location
# using by default kriging for prediction, and squared residual for non-conformity measure.
# s0 prediction location(s)
# s coordinates of locations.
# Y 
# eta kernel bandwidth
# m num of nearest locations to use for prediction.
# thetaHat [nugget, sill, range, smoothness[] from get_theta
# precision 
# alpha significance level. Defaults to 0.05.
# return lower and upper pred bounds at s0.

scp <- function(s0,s,Y,eta=Inf,m=NULL,thetaHat=NULL,precision=NULL,alpha=0.05){

  gamma <- data.frame(lower = NA, upper = NA)
  
  .prime(s0,s,Y,eta,m) # output Y_aug w_aug d_aug
  Q <- solve(mat_cov(d_aug,thetaHat))
  
  p_df  <- .plausibility_contour(Y_aug,w_aug,Q) 
  Y_cand <- p_df$Y_cand
  p_y  <- p_df$p_y

  gamma <- range(Y_cand[which(p_y >= floor((M+1) * alpha) / (M+1))])

  return(gamma)
}
```


```{r, eval=FALSE}
############################## test scp

Y[61,] # 1.799352  at (0.5, 0.5)

scp.test <- scp(s0=s0,s=s[-61,],Y=Y[-61], m=20, thetaHat=thetaHat, precision=0.5)

scp.test # -26.91262  25.81170

scp.test <- scp(s0=s0,s=s[-61,],Y=Y[-61], eta=0.3, m=20, thetaHat=thetaHat, precision=0.5)

scp.test # change eta, same results......

```

